---
title: "The Office Lines Analysis"
description: |
  Analysis of The Office Lines data from Tidy Tuesday (https://www.kaggle.com/datasets/fabriziocominetti/the-office-lines?resource=download)
author: Kandace Loualhati
date: September 30, 2024
format: html
execute:
  warning: false
  message: false
---

```{r}
# load in packages
library(tidyverse)
library(readr)
library(dplyr)
library(knitr)

# import data set
office_lines <- read.csv("the-office_lines.csv")

# filter out Jim's lines when he says "Pam" in every season
Jim_lines <- office_lines |>
  filter(Character == "Jim") |>
  filter(str_detect(Line, "\\bPam[a-z]*\\b")) |>
  group_by(Season, Episode_Number) |>
  summarize(n = n(), .groups = "drop") 

# plot of Jim's frequency of the word "Pam"
ggplot(Jim_lines, aes(x = Episode_Number, y = n, fill = Season)) +
  geom_col() +
  labs(
    x = "Episode Number",
    y = "# of 'Pam' Occurrences", 
    title = "How Many Times Jim Said 'Pam' in 'The Office'"
  )
```

The Office, although heavily referred to as a "comedy," has one of the most popular romantic couples in television sitcom history: Jim and Pam. In this analysis, I counted how many times Jim said the name "Pam" / "Pamela" per episode in every season. The graph is also color-coded by season: getting a lighter shade of blue as the seasons progress. Based on the data, Jim said the name "Pam" the most on episode 4 collectively throughout every season, and this trend would decrease the longer a season went on. Ultimately, Jim referenced "Pam" in 123 episodes throughout the course of the TV show.

```{r}
# detect whenever somebody laughs
laugh_counts <- office_lines |>
  filter(str_detect
         (str_to_lower
           (str_trim(Line)), "\\[laughing\\]|\\[laughs\\]")) |>
  group_by(Episode_Number, Season) |>
  summarize(n = n()) |>
  arrange(desc(n))

# create a table indicating laugh counts
kable(laugh_counts, caption = "Summary of Laugh Counts by Character and Episode")
```
